---
title: "Exercises Session 06-11-2019"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise Session 06-11-2019

#Group: Santiago Blanco Sanchez, Lais Silva Almeida Zanchetta, Daniel Ferreira Zanchetta

## Exercises Block I

### 1)
```{r}
vector<-c(1:20)
vectorWithPos2 <- vector[2]
vector[2]<-vector[20]
vector[20]<-vectorWithPos2
vector
```

### 2)
Step 1: Create the vector
```{r}
a <- rep(c(1,2,3,NA),5)
```
Step 2: Find out the mean, without considering the NA value
```{r}
mean <- mean(a,na.rm=TRUE)
```
Step 3: Create one recursive function that will be used later onto sapply
```{r}
fun <- function(x){ if(is.na(x)){x<-mean} else{x<-x}}
```
Step 4: Assign the sapply function to the vector at step 1
```{r}
a<-sapply(a, fun)
a
```

### 3)
First, create 3 vectors and, then, apply the data.frame function.
```{r}
Age <- c(22,25,18,20)
Name <- c("James","Mathew","Olivia","Stella")
Gender <- c("M","M","F","F")
df <- data.frame(Age,Name,Gender)
df
```

### 4)
```{r}
df[Age>21,]
```

### 5)
```{r}
df["adult"]<-c(df$Age>21)
df
```

### 6)
Instructions to write the iris.csv file and read it:
```{r}
data("iris")
write.table(iris,file="iris.csv",row.names=FALSE, col.names=TRUE,sep=";")
irisdf <- read.table("iris.csv",header=TRUE,sep=";")
```
Second set of instruction to calculate by rows and columns
```{r}
iris_subset <- unlist(lapply(irisdf, is.numeric))
```
```{r}
#Using apply
apply(irisdf[,iris_subset], 1, mean)
apply(irisdf[,iris_subset], 2, mean)
```
```{r}
#Using colMeans and rowMeans
colMeans(irisdf[ , iris_subset])
rowMeans(irisdf[ ,iris_subset])
```

### 7)
This function should consist of the following steps:
1)Define a subset of the iris dataframe, to contain only numeric values. Otherwise, it wont be possible to apply the rowMeans and colMeans functions (both expect numeric dataframe/matrix/vector/list)
2)Calculate the Row Means
3)Calculate the Column Means
4) Return both values
```{r}
funMean <- function(dataset){
  subsetDataSetNumeric <- unlist(sapply(dataset,is.numeric))
  print(rowMeans(dataset[,subsetDataSetNumeric]))
  print(colMeans(dataset[,subsetDataSetNumeric]))
}
funMean(irisdf)
```

### 8)
```{r}
funFibonacci <- function(n){
  x <- numeric(n)
  n1 <- -1
  n2 <- 1
  for(i in 1:length(x)){
   x[i] <- n1 + n2
   n1 <- n2
   n2 <- x[i]
  }
  x
}
funFibonacci(10)
```

### 9)
Firstly reading the dataset
```{r}
precipitaciones <- read.table("C:/Users/Daniel/Documents/precipitacionsbarcelonadesde1786.csv", header=TRUE, sep=",")
View(precipitaciones)
```
Secondly, some of the functions seen have been applied to the dataset
```{r}
colSums(precipitaciones,na.rm=TRUE)
rowSums(precipitaciones[,2:13],na.rm=TRUE)
colMeans(precipitaciones, na.rm=TRUE)
rowMeans(precipitaciones[,2:13],na.rm=TRUE)
```
Finally, exploring the plots
```{r}
y <- rowMeans(precipitaciones[,2:13],na.rm=TRUE)
x <- precipitaciones[1]
dataToPlot <- data.frame(x,y)
plot(dataToPlot,type="p",xlab="Year",ylab="Average")
hist(precipitaciones$Precip_Acum_Novembre, col="blue")
```

## Exercises Block II

###1)
```{r}
choices<-read.table("C:\\Users\\Daniel\\Documents\\Certificados & Faculdade\\UPC Master Big Data\\Data Analytics\\Choices.csv", header=TRUE, sep =";")
View(choices)
str(choices) #Code to show the class of all variables
```

###2)
```{r}
choices_conting<-table(choices$INFO,choices$MEASURE) #creation of the contingency table
choices_conting
```

###3)
```{r}
conTableWithSums<-data.frame(addmargins(choices_conting))
conTableWithSums
```

###4)
```{r}
chisq.test(choices_conting)
```
Reply: In this scenario, given the p-value less than the 0.05 level, we can undertand that Measure and Info are dependent variables.